Holomorphic field theories and higher algebra

### 2024-10-01 draft.tex
more small changes

In this subsection we discuss some global aspects of holomorphic factorization algebras.

A map $f \colon X \to Y$ induces a pushforward between categories of factorization algebras.
If $\cF$ is a factorization algebra on $X$, then the pushforward $f_* \cF$ is the factorization algebra on $Y$ which assigns to an open set $U \subset Y$ the value $\cF(f^{-1}(U))$.
Suppose we are in the situation where $\cF$ is the factorization algebra of observables of some quantum field theory on $X$.
In the case that $f$ is a smooth map with compact fibers, then $f_* \cF$ is the factorization algebra of observables for the \textit{compactification} of the quantum field theory along $f$.
In the case $f \colon Z \times Y \to Y$ is the projection, one says that $f_* \cF$ is the compactification of $\cF$ along $Z$.

Another special case of this is when $f$ is the map $X \to \star$.
Then $f_* \cF = \int_X \cF$ is the global sections of the factorization algebra $\cF$ along $X$.
In the case that $\cF$ is a locally constant factorization algebra this agrees with the factorization homology of $\cF$ along $X$.
In the case that $\cF$ is a holomorphic factorization algebra on a Riemann surface $\Sigma$, the object $\int_\Sigma \cF$ is closely related to the chiral homology (or conformal blocks) of the chiral (or vertex) algebra associated to $\cF$ along $\Sigma$.

In this way, when $X$ is a general complex manifold and $\cF$ is a holomorphic factorization algebra, then $\int_X \cF$ is a natural higher-dimensional generalization of the notion of conformal blocks of $\cF$.

For complex manifolds, and holomorphic factorization algebras on them, we consider an explicit example that generalizes a well-known concept in complex dimension one.
First, consider a \textit{real} one-dimensional \textit{topological} factorization algebra $\cA$.
If we identify $\cA$ with the associative algebra that it determines, then the factorization homology $\int_{S^1} \cA$ along the circle agrees with the Hochschild cohomology of the associative algebra $\cA$.
In other words, it is the home for all of the `traces' which the algebra $\cA$ admits.
It is well-known that the elliptic curve, or genus one, conformal blocks of a vertex algebra $\mathbb{V}$ is the natural `home' for the characters of the vertex algebra.
The most basic character of a conformal vertex algebra is the $q$-character defined in terms of a trace:
\beqn
\op{char}_{\mathbb{V}} (q) = \op{Tr}_{\mathbb{V}} (q^{L_0-c/24}) ,
\eeqn
where $L_0$ denotes the action of the vector field $z \del_z$.
The shift by the central charge guarantees that for $\mathbb{V}$ conformal that this expression has nice modularity properties.
From the point of view of factorization algebras, we can understand this trace in terms of Hochschild homology of the \textit{modes} algebra associated to $\mathbb{V}$ as introduced in \S\ref{??}.
Indeed, suppose we know that $\mathbb{V}$ corresponds to a holomorphic factorization algebra $\cF_{\mathbb{V}}$ on $\C$.
Then, if $r \colon \C^\times \to \RR_{>0}$ is the radial projection, we have discussed how to extract the modes algebra from the one-dimensional factorization algebra $r_* \cF_{\mathbb{F}}$.
If we further put this on the circle, we find Hochschild homology.
Thus, for every modulus $q$ one has
\beqn
\int_{E_q} \cF_{\mathbb{V}} = \int_{S^1} r_* \cF_{\mathbb{V}} = \op{Hoch}(r_* \cF_{\mathbb{V}}) .
\eeqn
For more on this perspective see \cite{GuiLi}.
From this point of view, the factorization homology along an elliptic curve encodes characters in the sense of vertex algebras.

In higher dimensions we can build a similar picture.
The simplest so-called \textit{Hopf manifolds} are complex manifolds of the form
\beqn
H_{\mathbf{q}} = \left(\C^d - \{0\}\right) \slash \left((z_1,\ldots,z_d) \sim (q_1 z_1,\ldots,q_d z_d) \right)
\eeqn
where $\mathbf{q} = (q_1,\ldots,q_d)$ are complex numbers satisfying $0 < |q_i| < 1$.
For generic $\mathbf{q}$ such a complex manifold is diffeomorphic to $S^{2d-1} \times S^1$.
The factorization homology of a holomorphic factorization algebra $\cF$ along a Hopf surface of this kind can be identified with the Hochschild homology of the associative algebra of $S^{2n-1}$-modes:
\beqn
\int_{H_{\mathbf{q}}} \cF \simeq \op{Hoch}(r_* \cF)
\eeqn
where, again, $r\colon \C^d - \{0\} \to \RR_{>0}$ is the radius.
Thus, we see that the factorization homology $\int_{H_{\mathbf{q}}} \cF$ naturally encodes traces of the $S^{2n-1}$-mode algebra associated to $\cF$.
In \cite{SWchar}, it is shown how an important invariant called the \textit{supersymmetric index} (or supersymmetric partition function) of a supersymmetric theory along a sphere $S^{2n-1}$ be cast into this framework at the level the holomorphic twist.

@@

In other words, divergences do not appear.

@@

The discussion so far concerned the analytic issues that are present when trying to make the path integral precise.
For some quantum field theories, like gauge theories, there is another issue that could prevent the theory from being well-posed:
these are called {\em anomalies}.
An anomaly measures the failure of the path integral to enjoy the same symmetries as the underlying classical field theory.
In gauge theory, for instance, the perturbative renormalization might not be gauge invariant, although the action functional is.
In topological theories, it is known that these sort of anomalies are absent.
\owen{I am confused by this comment. Aren't there TFTs with anomalies of some sort?}
For holomorphic theories, however, anomalies are present.\footnote{Anomalies in holomorphic field theory are related to, but distinct from, anomalies well-known to physicists, such as ``chiral anomalies'' in theories of Weyl fermions.}
Since renormalization is so well-behaved, however, such anomalies are amenable to explicit characterization.
For example, at one-loop in perturbation theory, an explicit formula for such anomalies for holomorphic theories on $\C^d$ is given in \cite{Whol}.
It involves a sum over Feynman weights $W_\Gamma$ where $\Gamma$ is a polygon with $(d+1)$~vertices.

>>>>

Factorization algebras yield interesting information when their global sections are computed,
much as sheaf cohomology is a rich source of invariants and information.
We focus here on some global aspects of holomorphic factorization algebras,
connecting with important results about vertex algebras and with the physical notion of compactification.

A map $f \colon X \to Y$ induces a pushforward functor between categories of factorization algebras.
If $\cF$ is a factorization algebra on $X$, 
then the pushforward $f_* \cF$ is the factorization algebra on $Y$ that assigns the value $\cF(f^{-1}(U))$ to an open set $U \subset Y$.

Consider a situation where $\cF$ is the factorization algebra of observables of a quantum field theory on $X$.
In the case that $f$ is a smooth map with compact fibers, 
then $f_* \cF$ is the factorization algebra of observables for the \textit{compactification} of the quantum field theory along $f$.
In the case $f \colon Z \times Y \to Y$ is the projection, 
one says that $f_* \cF$ is the compactification of $\cF$ along~$Z$.
\owen{We also used the terminology "dimensional reduction" earlier, so we should call back or distinguish these ideas.}

Another special case is the map $f: X \to \star$.
Then $f_* \cF = \int_X \cF$ is the global sections of the factorization algebra $\cF$ along $X$.
In the case that $\cF$ is a locally constant factorization algebra, 
this computation agrees with the factorization homology of $\cF$ along $X$, as defined by Ayala-Francis and Lurie \cite{}.
In the case that $\cF$ is a holomorphic factorization algebra on a Riemann surface $\Sigma$, 
the global sections $\int_\Sigma \cF$ is closely related to the chiral homology (or conformal blocks) of the chiral (or vertex) algebra associated to $\cF$ along $\Sigma$.
\owen{I want to assert "(It is the linear dual.)"}

Thus, it is natural to view global sections as a higher-dimensional generalization of the notion of conformal blocks.
That is, when $X$ is any complex manifold and $\cF$ is a holomorphic factorization algebra on $X$, 
then $\int_X \cF$ plays the role of conformal blocks.
One can then ask for analogs in higher dimensions of important results involving Riemann surfaces and vertex algebras.
We sketch here an example of how the character of a (conformal) vertex algebra relates to the conformal blocks over the moduli of elliptic curves.

First, consider a \textit{real} one-dimensional \textit{topological} factorization algebra $\cA$.
If we identify $\cA$ with the associative algebra that it determines, 
then the factorization homology $\int_{S^1} \cA$ along the circle agrees with the Hochschild cohomology of the associative algebra~$\cA$.
In other words, it is the home for all of the ``traces'' defined on the algebra~$\cA$.

For a vertex algebra $\mathbb{V}$, the role of the circle is played by any elliptic curve,
and the role of Hochschild homology is played by conformal blocks on that genus one curve. 
Thus conformal blocks over elliptic curves provides the natural home for the characters of the vertex algebra.
The most basic example of the character of a conformal vertex algebra is its $q$-character,
which is defined in terms of a trace:
\beqn
\op{char}_{\mathbb{V}} (q) = \op{Tr}_{\mathbb{V}} (q^{L_0-c/24}) ,
\eeqn
where $L_0$ denotes the action of the vector field $z \del_z$.
The shift by the central charge guarantees that this expression has nice modularity properties, i.e., we can view $q$ as parametrizing the moduli space of elliptic curves.
\owen{Is that an ok turn of phrase?}

From the point of view of factorization algebras, we can understand this trace in terms of Hochschild homology of the \textit{modes} algebra associated to $\mathbb{V}$ as introduced in \S\ref{??}.
Indeed, suppose the conformal vertex algebra $\mathbb{V}$ arises from a holomorphic factorization algebra $\cF_{\mathbb{V}}$ on $\C$.
Then, if $r \colon \C^\times \to \RR_{>0}$ is the radial projection, we have discussed how to extract the modes algebra from the one-dimensional factorization algebra $r_* \cF_{\mathbb{F}}$.
If we further put this on the circle, we find Hochschild homology.
Thus, for every modulus $q$ one has
\beqn
\int_{E_q} \cF_{\mathbb{V}} = \int_{S^1} r_* \cF_{\mathbb{V}} = \op{Hoch}(r_* \cF_{\mathbb{V}}) .
\eeqn
\owen{I think $\int_{E_q}$ receives a map from $\op{Hoch}(r_* \cF_{\mathbb{V}})$ but I don't see why it's an isomorphism. On the other hand, given a conformal block (i.e., linear functional on $\int_{E_q}$, we can compose it with that map to define a trace on the modes algebra,
which ought to be the character.}
For more on this perspective see \cite{GuiLi}.
From this point of view, the factorization homology along an elliptic curve encodes characters in the sense of vertex algebras.

In higher dimensions we can build a similar picture,
replacing the moduli of elliptic curves with the moduli of {\it Hopf manifolds}.
Recall that a Hopf manifold is a quotient of the punctured affine plane
\beqn
H_{\mathbf{q}} = \left(\C^d - \{0\}\right) \slash \left((z_1,\ldots,z_d) \sim (q_1 z_1,\ldots,q_d z_d) \right)
\eeqn
where $\mathbf{q} = (q_1,\ldots,q_d)$ are complex numbers satisfying $0 < |q_i| < 1$.
For generic $\mathbf{q}$ such a complex manifold is diffeomorphic to $S^{2d-1} \times S^1$.
The factorization homology of a holomorphic factorization algebra $\cF$ along a Hopf surface of this kind can be identified with the Hochschild homology of the associative algebra of $S^{2n-1}$-modes:
\beqn
\int_{H_{\mathbf{q}}} \cF \simeq \op{Hoch}(r_* \cF)
\eeqn
where, again, $r\colon \C^d - \{0\} \to \RR_{>0}$ is the radius.
Thus, we see that the factorization homology $\int_{H_{\mathbf{q}}} \cF$ naturally encodes traces of the $S^{2n-1}$-mode algebra associated to $\cF$.
In \cite{SWchar}, it is shown how an important invariant called the \textit{supersymmetric index} (or supersymmetric partition function) of a supersymmetric theory along a sphere $S^{2n-1}$ be cast into this framework at the level the holomorphic twist.

@@

In other words, divergences do not appear when constructing the perturbative quantization of {\em any} holomorphic field theory on~$\CC^d$.

@@

The discussion so far concerned the analytic issues that are present when trying to make the path integral precise.
For some quantum field theories, like gauge theories, there is another issue that could prevent the theory from being well-posed:
these are called {\em anomalies}.
An anomaly measures the failure of the path integral to enjoy the same symmetries as the underlying classical field theory.
In gauge theory, for instance, the perturbative renormalization might not be gauge invariant, although the action functional is.
In topological theories, it is known that these sort of anomalies are absent.
\owen{I am confused by this comment. Aren't there TFTs with anomalies of some sort?}
For holomorphic theories, however, anomalies are present.\footnote{Anomalies in holomorphic field theory are related to, but distinct from, anomalies well-known to physicists, such as ``chiral anomalies'' in theories of Weyl fermions.}
(Within the context of Costello's formalism for perturbative QFT,
these anomalies arise as obstructions to solving a Maurer-Cartan equation known as the quantum master equation.)

Since renormalization is so well-behaved, however, such anomalies are amenable to explicit characterization.
For example, at one-loop in perturbation theory, an explicit formula for such anomalies for holomorphic theories on $\C^d$ is given in \cite{Whol}.
It involves a sum over Feynman weights $W_\Gamma$ where $\Gamma$ is a polygon with $(d+1)$~vertices.
